Boosting Higgs Pair Production in the $bb\bar{b}\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-07-08 paper/sec-intro.tex
work on the paper

In the Standard Model (SM), the dominant mechanism for the production of two Higgs bosons at the LHC is 
gluon fusion (see Ref.~\cite{baglio} and references therein), analogously to single Higgs production.
For a center-of-mass energy $\sqrt{s} = 14\,$TeV, the recently computed next-to-next to
leading order (NNLO)
total cross section is approximately $40\,$fb~\cite{deFlorian:2013jea}.
%
Feasibility studies in the case of a light Higgs boson have been performed for several different final states, including
$b\bar b\gamma\gamma$~\cite{Baur:2003gp,Barger:2013jfa},
$b\bar{b}\tau^+\tau^-$~\cite{Baur:2003gpa,Barr:2013tda,Dolan:2012rv,Dolan:2013rja},
$b\bar{b}W^+W^-$~\cite{Dolan:2012rv,Papaefstathiou:2012qe} and
$b\bar{b}b\bar{b}$~\cite{Baur:2003gpa,Dolan:2012rv,Gouzevitch:2013qca,Cooper:2013kia,Wardrope:2014kya,deLima:2014dta}.
%
While these various studies differ in their qualitative conclusions, it is clear
that the ultimate precision in the determination of the Higgs trilinear
will be provided by the combination of various final states.
%

The advantage of the $b\bar{b}b\bar{b}$ final state is the the signal yield
is enhanced by the large branching fraction of Higgs bosons into bottom-anti-bottom
pairs, ${\rm BR}\lp H\to b\bar{b}\rp\simeq 0.57$~\cite{Dittmaier:2012vm}.
%
However, this channel needs to deal with an overwhelming QCD multi-jet background,
which makes this measurement very challenging.
%
Previous studies of Higgs pair production in this
channel~\cite{Wardrope:2014kya,deLima:2014dta}, considering
only the $4b$ irreducible QCD background, estimate that with a luminosity of
3000 fb$^{-1}$ a signal significance of $S/\sqrt{B}=$AAA can be obtained.
%
However, as show in this work, neglecting other QCD backgrounds such as $bbjj$ and
$jjjj$, with some of these light jets mistagged as $b$-jets, cannot be neglected, making
this measurement even more difficult.

Despite these large backgrounds, by combining the contribution of different final state
topologies, resolved, boosted and intermediate, with a careful optimization
of the analysis strategy, we find that it is possible to obtain a reasonable
good signal discrimination.
%
To further enhance the significance of the measurements, signal and background events
are processed using a Multivariate Analysis (in particular a feed-forward neural
network), that further enhance $S\sqrt{B}$ up to a value of AAA.
%
Therefore, we are able to rescue the potential of this final state in a realistic
analysis setup.

@@

The structure of this paper is the following.
%
In Sect.~\ref{mcgeneration} we discuss the generation of the signal
and background Monte Carlo samples.
%
Then in Sect.~\ref{sec:analysis}
we describe our analysis strategy, and in particular
the classification of each individual event into
different categories according to the event topology.
%
The results for our traditional cut-based analysis
are presented in Sect.~\ref{sec:results}, and in
Sect.~\ref{sec:mva} we show how we can enhance signal
significance by means of multivariate techniques.
%
Finally in Sect.~\ref{sec:conclusions} we conclude and outline
possible future developements for Higgs pair production
in the $b\bar{b}b\bar{b}$ final state.

>>>>

In the Standard Model (SM), the dominant mechanism for the production of two Higgs bosons at the LHC is 
gluon fusion (see Ref.~\cite{baglio} and references therein), analogously to single Higgs production.
For a center-of-mass energy $\sqrt{s} = 14\,$TeV, the recently computed next-to-next to
leading order (NNLO)
total cross section is approximately $40\,$fb~\cite{deFlorian:2013jea}.
%
Feasibility studies in the case of a light Higgs boson have been performed for several different final states, including
$b\bar b\gamma\gamma$~\cite{Baur:2003gp,Barger:2013jfa},
$b\bar{b}\tau^+\tau^-$~\cite{Baur:2003gpa,Barr:2013tda,Dolan:2012rv,Dolan:2013rja},
$b\bar{b}W^+W^-$~\cite{Dolan:2012rv,Papaefstathiou:2012qe} and
$b\bar{b}b\bar{b}$~\cite{Baur:2003gpa,Dolan:2012rv,Gouzevitch:2013qca,Cooper:2013kia,Wardrope:2014kya,deLima:2014dta}.
%
While these various studies differ in their qualitative conclusions, it is clear
that the ultimate precision in the determination of the Higgs trilinear
will be provided by the combination of various final states.
%
Remarkably, searches for the production of Higgs pairs, whose rates
could be enhanced by BSM effects have already been performed
by ATLAS~\cite{Aad:2015uka} and CMS~\cite{Khachatryan:2015yea},
though the sensitivity of these searches is still far from
that required for a SM production mechanism.

The advantage of the $b\bar{b}b\bar{b}$ final state is the the signal yield
is enhanced by the large branching fraction of Higgs bosons into bottom-anti-bottom
pairs, ${\rm BR}\lp H\to b\bar{b}\rp\simeq 0.57$~\cite{Dittmaier:2012vm}.
%
However, this channel needs to deal with an overwhelming QCD multi-jet background,
which makes this measurement very challenging.
%
Previous studies of Higgs pair production in this
channel~\cite{Wardrope:2014kya,deLima:2014dta}, considering
only the $4b$ irreducible QCD background, estimate that with a luminosity of
3000 fb$^{-1}$ a signal significance of around $S/\sqrt{B}=2.0$ can be obtained.
%
Both studies were based on exploiting the differences between signal and background
using techniques that account for a large number of variables, Boosted Decision Trees
and Shower Reconstruction respectively.
%
However, as show in this work, neglecting other QCD backgrounds such as $bbjj$ and
$jjjj$, with some of these light jets mistagged as $b$-jets, cannot be neglected, making
this measurement even more difficult.

Despite these large backgrounds, by combining the contribution of different final state
topologies, resolved, boosted and intermediate, with a careful optimization
of the analysis strategy, we find that it is possible to obtain a reasonable
good signal discrimination.
%
To further enhance the significance of the measurements, signal and background events
are processed using a Multivariate Analysis (in particular a feed-forward neural
network), that further enhance $S/\sqrt{B}$ up to a value of 3.0
%
Therefore, we are able to rescue the potential of this final state in a realistic
analysis setup.

@@

The structure of this paper is the following.
%
In Sect.~\ref{mcgeneration} we discuss the generation of the signal
and background Monte Carlo samples.
%
Then in Sect.~\ref{sec:analysis}
we describe our analysis strategy, and in particular
the classification of each individual event into
different categories according to the event topology.
%
The results for our traditional cut-based analysis
are presented in Sect.~\ref{sec:results}, and in
Sect.~\ref{sec:mva} we show how we can enhance signal
significance by means of multivariate techniques.
%
Finally in Sect.~\ref{sec:conclusions} we conclude and outline
possible future developments for Higgs pair production
in the $b\bar{b}b\bar{b}$ final state.

### 2015-07-08 paper/sec-mva.tex
work on the paper

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \vspace{-1cm}
  \includegraphics[width=0.80\textwidth]{plots/res_nnarch.pdf}
  \vspace{-1cm}
\caption{\small Architecture of the ANN used for the analysis of the resolved
  category, with $N_{\rm var}=13$ input variables and thus the same number of neurons
in the first layer.}
\label{fig:nnarch}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

\item The ratio of 2-to-1-Subjetiness $\tau_{12}$~\cite{Thaler:2010tr,Thaler:2011gf}.

@@

\item The energy correlation function double-ratios $C_2$ and $D_2$~\cite{Larkoski:2013eya}.

  The energy correlation double ratio $C_2$ is defined
  as
  \be
  \label{eq:c2}
C_2^{\beta} \equiv \frac{ {\rm ECF}(3,\beta) {\rm ECF}(1,\beta)}{{\rm ECF}(2,\beta)} \, ,
  \ee
  where the energy correlation functions ${\rm ECF}(3,\beta)$ are defined
  in~\cite{Larkoski:2013eya} with the motivation that $(N+1)$-point correlators
  are sensitive to $N$-prong substructure.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \vspace{-1cm}
  \includegraphics[width=0.90\textwidth]{plots/res_nnarch.pdf}
  \vspace{-1cm}
\caption{\small Architecture of the ANN used for the analysis of the resolved
  category, with $N_{\rm var}=13$ input variables and thus the same number of neurons
  in the first layer.
  %
  The color code in the neuron connections (the weights) is a heat map,
  with red indicating large values and black small values.
}
\label{fig:nnarch}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

\item The ratio of 2-to-1-subjettiness $\tau_{12}$~\cite{Thaler:2010tr,Thaler:2011gf}.

@@

\item The energy correlation function double-ratios $C_2$ and $D_2$~\cite{Larkoski:2013eya,Larkoski:2014gra}.

  The energy correlation double ratios  $C_2^{\beta}$
  and $D_2^{\beta}$ are defined
  as follows
  \be
  \label{eq:c2}
C_2^{\beta} \equiv \frac{ {\rm ECF}(3,\beta) {\rm ECF}(1,\beta)}{\lc {\rm ECF}(2,\beta)\rc ^2} \, ,
\ee
 \be
  \label{eq:d2}
D_2^{\beta} \equiv \frac{ {\rm ECF}(3,\beta)}{\lc {\rm ECF}(2,\beta)\rc^3} \, ,
  \ee
  where the energy correlation functions ${\rm ECF}(N,\beta)$ are defined
  in~\cite{Larkoski:2013eya} with the motivation that $(N+1)$-point correlators
  are sensitive to $N$-prong substructure.
  %
  The free parameter $\beta$ is set to a value of $\beta=2$,
  as recommended by the authors of~\cite{Larkoski:2013eya,Larkoski:2014gra}.

@@

Finally, let us briefly comment on our results would be affected
if we take into account the systematic uncertainty in the measurement
of the backgrounds, in addition to the statistical uncertainty.
%
Assuming an optimistic value of a total 3\% systematic uncertainty in the
multijet background measurement, we find that the signal significance is
modified as....
{\bf complete}
